---
title: "Intro to API with R"
author: "Mike J"
date: "5/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduciton to API's with R

Set of programming instructions for accessing information or services from a Web-based software application. 

Companies releases its API to the public so that other software developers can design products that can access its services.


<br>

Good website that shows information on available APIs and 
how to access them.

https://www.springboard.com/library/data-science/top-apis-for-data-scientists/

<br>
<br>
##How can API help data analysts
Can periodically analyze data without time consuming process of preparing new data tables manually.
Reduce human errors when constructing these tables.

<br>
<br>
<br>

##API KEY
Used to identify the calling program, its developer, or its user to the Web site.

Example: 842A3028-DFC5-43A4-9E38-94173687B03D

<br>
<br>
<br>

##httr and jsonlite packages

https://www.rdocumentation.org/packages/httr/versions/1.4.2/topics/GET

https://www.rdocumentation.org/packages/jsonlite/versions/1.7.2/topics/toJSON,%20fromJSON

<br>
<br>
```{r demo, echo=FALSE}
#install.packages(c("httr", "jsonlite"))

library(httr)
library(jsonlite)
library(dplyr)

#EIA data
# <EIA API KEY>

res <- GET('http://api.eia.gov/series/?series_id=ELEC.GEN.ALL-AK-99.A&api_key=<EIA API KEY>&out=json')

#check the status of the API request
httr::http_status(res)


```
```{r democontent, echo=FALSE}

httr::content(res)
vw<-rawToChar(res$content)
head(vw)

```


```{r demodata, echo=FALSE}


json_data <- fromJSON(rawToChar(res$content))
data <- data.frame(json_data$series$data)
data %>% rename(YEAR=X1, VALUE=X2)

```

```{r census, echo=FALSE}
#census key
#<CENSUS API KEY>

library(jsonlite)

#use the janitor package for data cleaning tasks that can't be done using dplyr
#install.packages("janitor")
library(janitor)

resURL <- "https://api.census.gov/data/2019/pep/charagegroups?get=NAME,POP&HISP=0&for=state:*"

ljson <- fromJSON(resURL)

#str(ljson )
#class(ljson )
head(ljson )

ljson2<-as.data.frame(ljson)
#using the janitor package

ljson3<-janitor::row_to_names(ljson2,1)
#str(ljson3 )
#class(ljson3 )
head(ljson3 )
```
<br>

##Some examples of packages that does the API request
<br>
<br>
```{r bls, echo=FALSE}
#Buearu of Labor Statistics data

library(blsAPI)
#library(rjson)


layoffs_json <- blsAPI('MLUMS00NN0001003') 
layoffs_df <- blsAPI('MLUMS00NN0001003', return_data_frame = TRUE)

layoffs_json
layoffs_df

```



```{r wb, echo=FALSE}
#install.packages("WDI")
library(WDI)

#search for series

WDIsearch(string='gdp', field='name', cache=NULL)

#pull data series

WDI(country=c("US","BR"), indicator="NY.GNS.ICTR.GN.ZS", start=1999, end=2000,
    extra=TRUE, cache=NULL)
```


```{r nooa, echo=FALSE}
#install.packages("rnoaa")
library(rnoaa)
options(noaakey = "<NOAA API KEY>")

ncdc(datasetid = 'GHCND', stationid = 'GHCND:USW00014895', startdate = '2013-10-01',
     enddate = '2013-12-01')
```


```{r fred,echo=FALSE}
#install.packages("fredr")
#library(dplyr)
#library(ggplot2)
library(fredr)

fredr_set_key("<FRED API KEY>")

popular_data_series<-fredr(
  series_id = "UNRATE",
  observation_start = as.Date("1990-01-01"),
  observation_end = as.Date("2000-01-01"),
  frequency = "q" # quarterly
) %>% 
ggplot(data = ., mapping = aes(x = date, y = value, color = series_id)) +
  geom_line() +
  labs(x = "Observation Date", y = "Rate", color = "Series")

popular_data_series
```

<br>
<br>
## twitteR package
twitteR is a popular service that allows users to broadcast short messages (’tweets’) for others to read. The twitteR package is intended to provide access to the Twitter API within R, allowing users to grab interesting subsets of Twitter data for their analyses.

<br>
<br>

```{r twitter, echo=FALSE}
#TWITTER
#install.packages("twitteR")
library(twitteR)

#load your information

api_key<-"<TWITTER API KEY>"

api_secret<-"<TWITTER API SECRET KEY>"

access_token<-"<TWITTER TOKEN KEY>"

access_token_secret<-"<TWITTER SECRET TOKEN KEY>"


setup_twitter_oauth(api_key,api_secret,access_token,access_token_secret)

#if your oauth function does not work then go to twitter and recreate your api key and token key.

```

<br>
<br>
<br>

```{r collectTwit, echo=FALSE}
obama_tweets<-searchTwitter("barackobama", n=8, lang="en")
df_obama=twListToDF(obama_tweets)
head(df_obama)

rstats_tweets<-searchTwitter("rstatstweet", n=8, lang="en")
df_rstats=twListToDF(rstats_tweets)
head(df_rstats)
```
<br>
<br>
##rtweet package:An implementation of calls designed to collect and organize Twitter data via Twitter's  Application Program Interfaces (API)
<br>
<br>

```{r rtweet, echo=FALSE}

#install.packages("rtweet")
library(rtweet)

app_name <- "<YOUR APP NAME>"
key <- "<TWITTER API KEY>"
secret <-"<TWITTER API SECRET KEY>"
token<-"<TWITTER TOKEN KEY>"
secreta<-"<TWITTER SECRET TOKEN KEY>" 

## create token

twitter_token <- create_token(
  app = app_name,
  consumer_key = key,
  consumer_secret = secret,
  access_token = token,
  access_secret = secreta)

rstats_tweets <- search_tweets(q = "#rstats",
                               n = 10)

rstats_tweets
```
<br>
<br>
##Lets Analyze our Tweets collected from Rstats
<br>
<br>

```{r clean, echo=FALSE}
library(tidytext)
library(ggplot2)
head(rstats_tweets$text)
#Data Clean-Up

rstats_tweets$stripped_text <- gsub("http.*","",  rstats_tweets$text)
rstats_tweets$stripped_text <- gsub("https.*","", rstats_tweets$stripped_text)

rstats_tweet_clean <- rstats_tweets %>%
  dplyr::select(stripped_text) %>%
  unnest_tokens(word, stripped_text)


```


```{r analysis, echo=FALSE}


# plot the top 15 words -- notice any issues?
rstats_tweet_clean %>%
  count(word, sort = TRUE) %>%
  top_n(15) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(x = "Count",
       y = "Unique words",
       title = "Count of unique words found in tweets")

```


